---
title: "Time Series Forcasting report"
author: "Kevork Sulahian"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r}
library(readxl)
library(forecast)

df <- read_xls('economy.xls', sheet='2011-2019 NACE 2')


df = df[4,]
df = df[-c(1,3)]
# rownames(df) = df[1]
df = df[-1]
df = t(df)
df[110] = "155770.7"
df[] <- sapply(df[],function(x) as.numeric(as.character(x)))
df = as.numeric(df)
# df= df * 1000000
df = df[-c(109:120)]
ts = ts(df, start = c(2011,1), frequency = c(12))
ts.plot(ts)

```


## Decomposing Time Series

Decomposing a time series means separating it into its constituent components, which are usually a trend component and an irregular component, and if it is a seasonal time series, a seasonal component.

### Decomposing Seasonal Data
A seasonal time series consists of a trend component, a seasonal component and an irregular component. Decomposing the time series means separating the time series into these three components: that is, estimating these three components.


```{r}
ts_components <- decompose(ts)
plot(ts_components)

```

```{r}
fit <- auto.arima(ts,max.p = 5,max.q = 5,max.P = 5,max.Q = 5,max.d = 3,seasonal = TRUE)

fit_forecast = forecast(fit,h=24)
plot(fit_forecast)
```


# Forecasting
These estimations are what could happened in 2020 and 2021 if Covid and the war had not impacted our economy

## Forecasting for GDP

### if Covid did not occur

GDP growth was estimated at 7.5% for 2020 

### if Covid and the war did not occur

GDP growth was estimated at 7.5% for 2021

## Forecasting for Total Industry

### if Covid did not occur

Total Industry growth was estimated at 7.4% for 2020 

### if Covid and the war did not occur

Total Industry growth was estimated at 6.0% for 2021

## Forecasting for Manufactring

### if Covid did not occur

Manufactring growth was estimated at 7.8% for 2020 

### if Covid and the war did not occur

Manufactring growth was estimated at 7.3% for 2021

## Forecasting for Service

### if Covid did not occur

Manufactring growth was estimated at 12.7% for 2020 

### if Covid and the war did not occur

Manufactring growth was estimated at 12.5% for 2021

## Forecasting for trade

### if Covid did not occur

Manufactring growth was estimated at 7.89% for 2020 

### if Covid and the war did not occur

Manufactring growth was estimated at 7.48% for 2021