---
title: '2021'
author: "Kevork Sulahian"
date: "3/12/2021"
output:
  html_document:
    df_print: paged
  word_document: default
  pdf_document: default
---
```{r, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo=FALSE) 
```

```{r, echo=FALSE, warning=FALSE}
library(readr)
library(tidyverse)
library(stringr)
library("sf")
library("ggplot2")
theme_set(theme_bw())
library("rnaturalearth")
library("rnaturalearthdata")
library(data.table)
library(kableExtra)
library(tm)
library(RColorBrewer)
library(wordcloud)
data <- read_csv("2018-03-06-2021-03-12-Caucasus_and_Central_Asia-Armenia.csv")
data <- data[data$year == 2021,]
dt <- data[c(1:5),c(5,8,10,19,26)]

dt %>%
  kbl() %>%
  kable_material(c("striped", "hover"))


```
<br>
Names of the available columns in the dataset
<br>
```{r}
colnames(data)

# length(unique(data$data_id))

# more than half of the data is the same?
## not in the 2021 version
data <- data %>%
  distinct(data_id, .keep_all = T)
```
<br>
Number of unique events that are specified in the data set
<br>
As seen there are only 3 types of events in 2021, while in previous years there were more
```{r}


unique(data$event_type)

# unique(data$sub_event_type)
```
<br>
The main actors of said events
<br>

```{r}
unique(data$actor1)

# unique(data$assoc_actor_1)
# 
# # unique(data$source_scale)
# # Well they fixed the lat and long
# 
# fix_cord <- function(test) {
#   test <- as.character(test)
#   test <- strsplit(test,"")
#   
#   for(i in 1:length(test)) {
#     temp = unlist(test[i])
#     temp1 = paste0(temp[1],temp[2],".")
#     temp2 <-paste(temp[3:length(temp)], collapse = "")
#     temp3 <- as.numeric(paste0(temp1,temp2))
#     test[i] = temp3
#   }
#   
#   test <- unlist(test)
#   return(test)
# }
# 
# data$latitude <- fix_cord(data$latitude)
# data$longitude <- fix_cord(data$longitude)


# first jab at map

sites <- st_as_sf(data.frame(longitude = data$longitude,
                             latitude = data$latitude),
                  coords = c("longitude", "latitude"),
                  crs = 4326, agr = "constant")
cord_dens <- setDT(data)[, .N, longitude]

data2 <- merge(x = data,y = cord_dens,by= "longitude")

min_max_norm <- function(x) {
  (x - min(x)) / (max(x) - min(x))
}

data2$N <- min_max_norm(data2$N)+ 3

world <- ne_countries(scale='medium',returnclass = 'sf')


countries <- subset(world, admin %in% c("Armenia"))
```
<br>
The plot below shows the types of the evnets and where they occur.
Note that many of the locations are approximated to the nearest city or village
<br>

```{r}
(Armenia <- ggplot(data = countries) +
    geom_sf(fill = "antiquewhite1") +
    geom_sf(data = sites, size = data2$N, shape = 16, aes(color=data$event_type)) +
    xlab("Longitude")+ ylab("Latitude") + labs(color = "Event Types"))

```
<br>
Number of the events that have occured 
<br>

```{r, echo=FALSE}

by_events <- data %>%
  group_by(event_type) %>%
  count(event_type) %>%
  arrange(desc(n))

ggplot(by_events, aes(x = event_type, y = n, fill = event_type)) +
  geom_bar(stat = 'identity') + 
  theme(axis.text.x = element_blank()) +
  labs(title = "Number of events in the region")


```
<br>
<br>

```{r, echo = FALSE}
unique_sources <- unique(str_subset(data$source,"^[^;]+$",negate = FALSE))
# make a df with unique source and count

# sum(str_detect(data$source,unique_sources[3]))

sources_count <- data.frame(
  source = unique_sources,
  n = 0
)

for (i in 1:length(unique_sources)) {
  sources_count$n[i] = sum(str_detect(data$source,unique_sources[i]))
}

sources_count <- sources_count[order(sources_count$n,decreasing = T),]

top_10_sources <- sources_count[c(1:10),]
```
<br>

```{r}
ggplot(top_10_sources, aes(x = source, y = n, fill = source)) +
  geom_bar(stat = 'identity') + 
  theme(axis.text.x = element_blank()) +
  labs(title = "sources")


```


```{r, echo = FALSE}

# without ministry
# top_10_sources2 <- sources_count[c(2:11),]
# 
# ggplot(top_10_sources2, aes(x = source, y = n, fill = source)) +
#   geom_bar(stat = 'identity') + 
#   theme(axis.text.x = element_blank()) +
#   labs(title = "sources without the goverment")
```

```{r, echo = FALSE}
data$event_date <- as.Date(data$event_date, format = '%d %B %Y')

data3 <- data
data3$month <- month(data$event_date)
data3$year <- year(data$event_date)

library(plotly)

data3$year <- as.factor(data3$year)
p <- ggplot(data3, aes(x = month)) +
  geom_bar(aes(y = ..count.., fill = event_type)) +
  scale_fill_brewer(palette = "Set3") +
  ylab("n") +
  ggtitle("Type of the events over months")

fig <- ggplotly(p)

fig

```
<br>
```{r}
text <- data$notes
vs <- VectorSource(text)
corpus <- VCorpus(vs)
corpus <- tm_map(corpus, removePunctuation)
dtm <- TermDocumentMatrix(corpus,
                          control = list(removeNumbers = T, removePunctuation = T,
                                         stopwords= T, stemming = T))
inspect(dtm)
dtm_mat <- as.matrix(dtm)
freqs <- rowSums(dtm_mat)
df_freqs <- data.frame(terms = rownames(dtm_mat),
                       freq = freqs, stringsAsFactors = F)
df_freqs <- df_freqs[order(df_freqs$freq, decreasing = T),]
head(df_freqs)

library(RColorBrewer)
df_freqs_10 <- df_freqs[1:10,]
ggplot(df_freqs_10, aes(x = reorder(terms, freq), y = freq)) +
  geom_bar(stat="identity", fill = brewer.pal(n = 10, name="Spectral")) +
  coord_flip() + labs(x = "Terms", y = "Frequency", title = "top 10 words")

set.seed(1)
library(wordcloud)

wordcloud(words = df_freqs$terms, freq = df_freqs$freq, min.freq = 10,
          max.words = 200, random.order = F,
          colors = brewer.pal(8, 'Dark2'))


tdm <- TermDocumentMatrix(corpus,
                          control = list(removeNumbers = T, removePunctuation = T,
                                         stopwords= T, stemming = T, 
                                         weighting = weightTfIdf))
tdm_mat <- as.matrix(tdm)
freqs <- rowSums(tdm_mat)
df_freqs <- data.frame(terms = rownames(tdm_mat),
                       freq = freqs, stringsAsFactors = F)
df_freqs <- df_freqs[order(df_freqs$freq, decreasing = T),]

set.seed(1)
```
<br>
```{r}
wordcloud(words = df_freqs$terms, freq = df_freqs$freq, min.freq = 1,
          max.words = 200, random.order = F,scale=c(4,.5),
          rot.per=.5,
          colors = brewer.pal(8, 'Dark2'))


```